% section 5.1.1: Data Access

\subsubsection{Data Access}

The data access framework is represented by several packages with
names beginning with \texttt{daf\_} (primarily \texttt{daf\_utils}
and \texttt{daf\_persistence}).  This framework, introduced initially
in DC2, provides mean for reading and writing data and manipulating
those in memory.  In DC3a, the framework was upgraded to improve
database access, security, and performance as well as enable easier
use by developers.  

In DC2, the {\tt DataProperty} class was used as a generic container
for named data--that is, metadata.  In DC3a, this was replaced with
the new {\tt PropertySet} class.  By simplifying the interface and
unifying concepts, it helped to reduce complexity for developers,
remove dependency issues, and bring {\tt Policy} and {\tt PropertySet}
into closer alignment.  In particular, there is no distinction between
a {\tt PropertySet} containing a single key/value pair and one
containing multiple name--value pairs.  While a {\tt DataProperty}
attempted to maintain the order of the properties within it, a {\tt
PropertySet} has no defined order except within the vector of values
for a given key.  The new interface allows hierarchical {\tt
PropertySet} values to be created more easily, using ``parent.child''
notation in the key name instead of requiring the creation of a
separate {\tt DataProperty} instance for the child.  

The model behind {\tt PropertySet} is key to how we handle metadata in
general.  One recurring question is to what extent a {\tt PropertySet}
is appropriate for handling FITS header keywords.  FITS has a notion
of an order {\it across different keywords} (which {\tt PropertySet} does not
currently support) by virtue of the data being stored serially in a
file.  Preservation of that order is not normally important for
consuming the data (that is, with a few exceptions order has no
semantic significance), and we don't currently require it in our
application.  Nevertheless, it may be useful for ensuring
``round-tripping'' (i.e. ensuring that data that hasn't logically
changed looks exactly the same when written out as when it was read
in).  We believe that we could add support for preserving this order
to the {\tt PropertySet} if and when it is needed.   A related issue
is one of preserving the comment that can be attached to each FITS
name-value pair; for this, we can consider supporting annotations to
each {\tt PropertySet} entry.  

The persistence framework saw a few additions.  New formatters for PSFs
and their underlying kernels and spatial functions were written,
allowing those classes to be persisted to Boost archive files and XML
files.  This was a non-trivial task due to the large number and
diversity of subclasses involved.  The ability to persist and restore
PSFs enhanced the debuggability of the pipeline.

The formatters for {\tt Source}s and {\tt DiaSource}s (actually {\tt
PersistableSourceVector}s and 
\hfil\break\texttt{PersistableDiaSourceVector}s) were also updated 
to match the new contents of those classes.  These formatters were at
the heart of inter-slice communication in DC3a as {\tt Source}s used
for WCS and PSF determination were communicated via the database so
that an entire CCD's information could be used for the calculations.

A central, consistent facility for processing {\tt LogicalLocation}
strings was provided, allowing information from an event, other {\tt
Clipboard} items, or a {\tt Policy} to be substituted into them with
optional formatting.  This feature was used extensively to generate
pathnames and database URLs.  It allowed stage policies to be
platform agnostic while the pipeline orchestration layer combined
with the pipeline harness to specify translations from
platform-independent form to platform-dependent form.

Database access was improved by implementing the {\tt DbStorage}
portability layer directly in terms of the MySQL C API instead of using
CORAL/SEAL as a second level of indirection.  This reduced the size and
complexity of the software stack without sacrificing portability.
Security was enhanced by improving the specification of database
authentication credentials and the deprecation of shared database
accounts.  The database interface now allows the limited ability to
execute raw SQL statements (without return values), allowing some
computations to be done efficiently entirely on the database server, and
it also allows the use of expressions and multi-part field names in
queries.

The {\tt DateTime} class that is used as a utility to deal with dates and
times throughout the LSST software stack had its interface significantly
revamped to improve the clarity of the timebase (UTC or TAI) used for
its arguments and results.  (The internal representation of times always
uses TAI.)  Conversions from {\tt DateTime} to and from ISO8601-format
strings were implemented.

