\begin{thebibliography}{}

\bibitem[Axelrod {\it et al.} (2008)]{dc2} Axelrod, T., R. Allsman,
  J. Kantor, J. Myers,  F. Pierfederici, S. Monkewicz, G. Daues,
  S. Pietrowicz, R. Plane, J. Becla, K-T. Lim, A. Becker, L. Jones,
  R. Owen, N. Silvestri, R. Lupton, J. Bartels 2008, ``LSST Data
  Challenge 2'', LSST Document Repository, Document-5356.

\bibitem[Borne {\it et al.} (2009)]{borne09} Borne, K. D., R. Laher, Z. Ivezic, and N. Hamam 2009,
   Petascale Object Classification of the LSST Event Stream, 
   Poster presented at the 213th AAS Meeting, Long Beach, CA, 4-8
   January 2009.

\bibitem[Bruzual \& Charlot (2003)]{bruzChar03} Bruzual, G, and
  S. Charlot 2003, \mnras, 344, 1000.

\bibitem[Debattista \textit{et al.} (2006)]{Debattista06}
  {{Debattista}, V.~P. and {Mayer}, L. and {Carollo}, C.~M. and  
    {Moore}, B. and {Wadsley}, J. and {Quinn}, T.} 2006, \apj, 645, 209.

\bibitem[Gamma \textit{et al.}(1995)]{GoF}Erich Gamma, Richard Helm, Ralph Johnson, and John Vlissides 1995,
Elements of Reusable Object-Oriented Software.

\bibitem[Juri\'c {\it et al.} (2008)][JuricIvezic08] {{Juri{\'c}}, M. and {Ivezi{\'c}}, {\v Z}. and {Brooks}, A. and 
	{Lupton}, R.~H. and {Schlegel}, D. and {Finkbeiner}, D. and 
	{Padmanabhan}, N. and {Bond}, N. and {Sesar}, B. and {Rockosi}, C.~M. and 
	{Knapp}, G.~R. and {Gunn}, J.~E. and {Sumi}, T. and {Schneider}, D.~P. and 
	{Barentine}, J.~C. and {Brewington}, H.~J. and {Brinkmann}, J. and 
	{Fukugita}, M. and {Harvanek}, M. and {Kleinman}, S.~J. and 
	{Krzesinski}, J. and {Long}, D. and {Neilsen}, Jr., E.~H. and 
	{Nitta}, A. and {Snedden}, S.~A. and {York}, D.~G.} 2008,
  \apj, 673, 864.

\bibitem[Kitzbichler and White (2007)]{KitzbichlerWhite07}
  {{Kitzbichler}, M.~G. and {White}, S.~D.~M.} 2007, \mnras, 376, 2.

\bibitem[Laher {\it et al.}(2008)]{laher08} Laher, Russ R., Deborah Levine, Vince Mannings, 
   Peregrine McGehee, Jeonghee Rho, Richard A. Shaw, and Jeff Kantor 2008,
   LSST Science Data Quality Analysis Subsystem Design,
   Oral presentation given at the ADASS XVIII Conference, Quebec City, 
   (http://spider.ipac.caltech.edu/staff/laher/lsst/O7A2.pdf)

\bibitem[Lupton et al., 2001]{photoADASS}
Lupton R.~H. et al., 2001 `The SDSS Imaging Pipelines' in
ASP Conf. Set. 238, Astronomical Data Analysis Software and Systems X.
   
\bibitem[Pier et al., 2003]{SDSSAstrom} Pier, J.R., Munn, J.A., Hindsley, R.B., Hennessy, G.S.,
Kent, S.M., Lupton, R.H., \& Ivezi\'c, \v{Z}. 2003, AJ, 125, 1559

\bibitem[Press et al., 2007]{PressNR}
  W.~H. Press,  S.~A. Teukolsky, W.~T. Vetterling, B.~P. Flannery.
  Numerical Recipes: the Art of Scientific Computing.
  2007, Cambridge University Press.
   
\end{thebibliography}
