% Section 6.5: WCS Estimation

\subsection{WCS Determination} \label{sec:wcs}

WCS, or World Coordinate System is a convention for mapping positions
on an astronomical image to coordinates in the sky. We require that
our WCS map our positions with an accuracy of better than 0.1", driven
by the need to subtract images from templates in the {\tt diff\_im}
stage. For DC3a we implemented a method of determining a WCS solution
for an image using the {\tt astrometry.net}\footnote{http://astrometry.net} package.

{\tt astrometry.net} takes as input a list of positions pixel
positions of sources in an image, and compares pattern of stellar
positions, or asterisms, with a template list of stars with known
right ascension and declination. It can be used to match images of
arbitary pixel scale, location on sky, and magnitude limit, provided
there is a sufficient overlap between the set of sources in the image,
and those in the template.

\subsubsection{Interface}
Because {\tt astrometry.net} is written in C, and does not have a
defined public API, we created a C++ wrapper to provide an object
oriented interface, which is called a {\tt
  meas::astrom::net::GlobalAstrometrySolution}, hereafter referrred to as
a GAS. A GAS is initialised with a set of template files, the names of
which are set via policy. The policy file also defines the equinox
(e.g equinox=2000) and coordinate system (e.g raDecSys=FK5) of the
template. These two parameters are necessary to fully define a WCS
solution, but are not included in the template files produced by {\tt
  astrometry.net}. The list of sources to be solved for is set using
the {\tt setStarlist()} function, which takes an {\tt
  lsst::afw:detection::SourceSet} as its argument.


The GAS takes various parameters which can be set to speed the match. 

\begin{description}
\item[{\tt getMinimumImageScale(), getMaximumImageScale}] If the plate scale (in arcseconds per pixel) of the image is known or can be constrained, the volume of parameter space that {\tt astrometry.net} needs to search can be dramatically reduced. If the image scale is exactly known, use {\tt setImageScaleArcsecPerPixel()}

\item[{\tt setParity()}] Astromical images are normally oriented with North up and East to the right, or rotated from the same. However, the image may be inverted, so that North is up when East is to the left. {\tt astrometry.net} searches for both possible configurations, unless the parity is specified. Specifying the parity can reduce the time taken to solve an image by a factor of two. Legal values for parity are \code{NORMAL\_PARITY}, \code{FLIPPED\_PARITY}, or \code{UNKNOWN\_PARITY}, the default.

\item[{\tt setNumBrightObjects()}] Although there may be many hundreds of sources on an image, typically a small fraction of these are necessary to solve an image. This function instructs {\tt astrometry.net} to only use the brightest $N$ stars when making a match. Using fewer stars generally results in a faster match, but using too few can result in the algorithm failing to return any match. Typically a value between approximately 40 and 80 works well. Higher values tend to significantly slow the match. 
\end{description}

\subsubsection{Timing}

{\tt astrometry.net} comes supplied with a large template based on the
USNO-b catalogue, but we found this template unsatisfactory for two
reasons. The sheer size of the catalog ($\sim$30\,Gb) meant that disk
I/O was prohibitively expensive. We also found that the centroids of
individual objects differed by up to 0.3" between the CFHT images and
the positions measured by USNO-b. Instead we created a smaller
template (2\,Mb) from images of the CFHT field created by Swarp, and
found that method worked to our satisfaction.

We measured the time taken to solve each ccd in the CFHT field
v704893-e0 (30 images). In each case, we put no constraints on the
solution other than to set the number of objects to use to find a
solution to 70. Solving blindly (i.e without using an initial guess at
the position) took, on average, 41\,ms $\pm$ 76\,ms. The worst case
time was 300ms. When a field fails to solve, {\tt astrometry.net}
takes quite some (10-20 seconds) to search every possible combination
of stars before giving up.

% \subsubsection{Success rate}
% {\tt astrometry.net} claims over 99\% success rate in solving $r$
% band images from the SDSS survey {\tt etc..}
%
% Ray:  It's not clear what the intended content here should be,
% particularly considering the problems we had with USNO-B.  

\subsubsection{Improvements to the WCS class}
Astronomical images generally contain some distortion, i.e the
rectangular grid of pixels in an image does not map onto a retangular
grid of right ascension and declination. The WCS standard defines
several functional forms to describe this distortion.  {\tt
  astrometry.net} instead describes distortion coefficients using the SIP
convention (Shupe et al. 2005; ASPC 347 491), which was developed and
used by the Spitzer Science Center. SIP defines a set of polynomials
used to map the positions of the distorted pixels into an undistorted
space where they can be correctly transformed using the WCS. {\tt afw}
adopted the SIP convention, and the WCS class was extended to use the
SIP polynomials when converting from pixel to sky coordinates.


\subsubsection{Future Work}
Templates have been produced for the 4 CFHT deep fields. However, the proceedure for producing these templates is quite involved, and needs to be simplified so that templates can be produced on demand for new fields.
